---
title: "Jon Lester June 28, 2014 Pitches vs. NY Yankeees"
author: "Danny Malter"
date: "December 23, 2014"
output: html_document
---

```{r}
setwd("~/pitchRx/Jon Lester")
```

Load in the required packages
```{r, warning=FALSE, message=FALSE}
require(mosaic)
require(ggplot2)
require(dplyr)
require(ggplot2)
require(RSQLite)
require(pitchRx)
require(knitr)
require(animation)
```

Create a pitchRx database between the date ranges
```{r, warning=FALSE, message=FALSE, results='hide'}
db20140628 <- src_sqlite("MLB20140628.sqlite3", create = TRUE)
scrape(start = "2014-06-28", end = "2014-06-28", connect = db20140628$con)
```

Join the location and names table into a new que table.  Select Jon Lester.
```{r}
locations <- select(tbl(db20140628, "pitch"), 
                    pitch_type, px, pz, des, num, gameday_link, inning)
names <- select(tbl(db20140628, "atbat"), pitcher_name, batter_name, 
                num, b_height, gameday_link, event, stand)
que <- inner_join(locations, filter(names, 
                                    pitcher_name == "Jon Lester"),
                  by = c("num", "gameday_link"))
pitchfx <- collect(que)
```

Create a new column with the full pitch type name.  Full pitch type is not to be used in strikeFX.
```{r}
pitchfx$pitch_type_full <- factor(pitchfx$pitch_type,
                  levels=c("FF", "FC", "SI", "CH", "CU"),
                  labels=c("Four-seam FB","Fastball (cutter)",
                           "Sinker","Changeup", "Curveball"))
```

Count of each type of pitch
```{r}
table(pitchfx$pitch_type_full)
```

Count of each type of pitch by batter stand and inning
```{r}
with(pitchfx, table(pitch_type_full, stand))
with(pitchfx, table(pitch_type_full, inning))
```

Count of pitch type by umpire call
```{r}
ball.strike <- subset(pitchfx, des == "Ball" | des == "Swinging Strike" | des == "Called Strike")
ggplot(ball.strike, aes(x = pitch_type_full, fill = des)) +
    geom_bar(position = "dodge") +
    guides(fill=guide_legend(title='Call')) +
    ggtitle("Count of Pitch Type") +
    xlab('Pitch Type') +
    ylab('Count')
```

Pitch type thrown to each type of batter
```{r, warning=FALSE}
# Overlay a strike zone
topKzone <- 3.5
botKzone <- 1.6
inKzone <- -0.95
outKzone <- 0.95
kZone <- data.frame(
  x=c(inKzone, inKzone, outKzone, outKzone, inKzone),
  y=c(botKzone, topKzone, topKzone, botKzone, botKzone)
)

print(ggplot(pitchfx, aes(px, pz, color=batter_name)) + geom_point() +
  geom_path(aes(x, y), data=kZone, lwd=2, col="red") +
  ylim(0, 5) + facet_wrap(~ batter_name, ncol=3))
```

Visualize Pitches
All pitches
```{r, warning=FALSE}
strikeFX(pitchfx, geom="tile", layer=facet_grid(.~stand))
```

Just Called Strikes
```{r, warning=FALSE}
lester.called.strikes <- subset(pitchfx, des == "Called Strike")
strikeFX(lester.called.strikes, geom="tile", layer=facet_grid(.~stand))
```

Just Balls
```{r, warning=FALSE}
lester.balls <- subset(pitchfx, des == "Ball")
strikeFX(lester.balls, geom="tile", layer=facet_grid(.~stand))
```

Probability of a strike
```{r, warning=FALSE, message=FALSE}
require(mgcv)
```

```{r, warning=FALSE}
noswing <- subset(pitchfx, des %in% c("Ball", "Called Strike"))
noswing$strike <- as.numeric(noswing$des %in% "Called Strike")
m1 <- bam(strike ~ s(px, pz, by=factor(stand)) +
          factor(stand), data=noswing, family = binomial(link='logit'))
strikeFX(noswing, model=m1, layer=facet_grid(.~stand))
```


```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
data <- scrape(start="2014-06-28", end="2014-06-28")
```


```{r, eval=FALSE, echo=FALSE, fig.show='animate', warning=FALSE}
names <- c("Jon Lester")
atbats <- subset(data$atbat, pitcher_name %in% names)
pitchFX <- plyr::join(atbats, data$pitch, by=c("num", "url"), type="inner")
pitch <- subset(pitchFX, pitch_type %in% c("FF", "FC", "SI", "CU", "CH"))
pitch.offspeed <- subset(pitchFX, pitch_type %in% c("CU", "CH"))

animateFX(pitch.fast, layer=list(facet_grid(pitcher_name~stand, labeller = label_both), theme_bw(), coord_equal()))
animateFX(pitch.offspeed, layer=list(facet_grid(pitcher_name~stand, labeller = label_both), theme_bw(), coord_equal()))
ani.options(interval = 0.05)
saveHTML({animateFX(pitch, layer=list(facet_grid(pitcher_name~stand, labeller = label_both), theme_bw(), coord_equal()))}, img.name="lester_pitches")

ani.options(interval = 0.05)
saveSWF({animateFX(pitch.fast, layer=list(facet_grid(pitcher_name~stand, labeller = label_both), theme_bw(), coord_equal()))}, img.name="lester_pitches")
```


