---
title: "Classifying MLB Pitch Types Using Neural Nets"
layout: post
comments: true
category: R
---


{% raw %}

# Classifying Jake Arrieta's Pitch Types #

Major League Baseball currently has high definition camera's capturing everything from the release point of the ball being thrown to the amount of horizontal spin on the ball.  Rather than having a person manually write down what each pitch type is, an algorithm does this automatically.  The goal of this project is to use MLB's classifications to create run a neural network of my own that will match that of MLB.

The training dataset being used are all of Jake Arrieta's pitches thrown in the 2016 regualar season and the model is validated on Game 2 of the 2016 world series in which Arrieta threw 98 pitches.


- Load in the required packages

```{r, warning=FALSE}
library(nnet)
library(caret)
library(pitchRx)
library(dplyr)
library(plyr)
library(RSQLite)
library(devtools)
```

The following script was used to scrape data from the 2016 season and filtered for Jake Arrieta's pitches.  A similar script was ran for only data on 10/26/16, the date of Game 2 from the 2016 World Series.

```{r db, eval=FALSE, echo=FALSE}
db2016 <- src_sqlite("MLB2015_All.sqlite3", create = FALSE)
scrape(start = "2016-04-01", end = "2016-10-01", connect = db2016$con)

locations <- select(tbl(db2016, "pitch"), 
                    pitch_type, start_speed, end_speed, num, gameday_link,
                    pfx_x, pfx_z, vx0, vy0, vz0, ax, ay, az, break_y, 
                    break_angle, break_length, spin_dir, spin_rate)
names <- select(tbl(db2016, "atbat"), pitcher_name, batter_name, 
                num, gameday_link, stand)
que <- inner_join(locations, filter(names, pitcher_name == "Jake Arrieta"),
                  by = c("num", "gameday_link"))

pitchfx <- collect(que, n=Inf)
pitchfx <- as.data.frame(pitchfx)
arrieta <- subset(pitchfx, select = -c(pitcher_name, batter_name, stand, num, gameday_link))
```

```{r import, eval=FALSE}
# import function for plotting nnet
#source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff04441270351#6c34f1a4684a5/nnet_plot_update.r')

# Read in 2016 season data and Game 2 of the World Series
arrieta <- read.csv('~/arrieta.csv')
validation <- read.csv('~/world_series.csv')

# Split the regular season data into a new training and testing set.
intrain <- createDataPartition(arrieta$pitch_type, p=0.7, list=FALSE)
myTrain <- arrieta[intrain,]
myTest <- arrieta[-intrain,]
head(myTrain)
```

## Neural Net from nnet Package ##

```{r train}
set.seed(353)
nn_mod1 <- nnet(pitch_type ~ ., data = myTrain, 
           size = 10, rang = 0.1, decay = 5e-04, 
           maxit = 1000)

test.pred <- predict(nn_mod1, newdata = myTest, type = "class")
table(test.pred, myTest$pitch_type)

# plot each model
plot.nnet(nn_mod1)

# summary of model
summary(nn_mod1)

# structure (15 inputs, 10 hidden layers, 5 outputs)
nn_mod1$n
```

```{r validate}
# Validation on new data from 2016 World Series game 2
val.pred <- predict(nn_mod1, newdata = validation, type = "class")
val.pred

# Misclassified Pitch number 13
table(val.pred, validation$pitch_type)
```

{% endraw %}

<script>
  (function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
  (i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
  m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
  })(window,document,'script','//www.google-analytics.com/analytics.js','ga');

  ga('create', 'UA-57468410-2', 'auto');
  ga('send', 'pageview');

</script>
